# Text Analysis I: Basics

## Goals{#goals07}

- introduce basic text analysis concepts:
  - ...
  - ...
  - ...

## Preliminaries{#Prelim07}

### Data{#Data07}

We will use the following text files in this worksheet. Please download them and keep them close to your worksheet. Since some of the files are quite large, you want to download them before loading them into R:

* [The Richmond Dispatch (1862)](./files/data/dispatch_1862.tsv)
* [Star Wars I: The Phantom Menace (script)](./files/data/sw1.md)

In order to make loading these files a little bit easier, you can paste the path to where you placed these files into an isolated variable and then reuse it as follows (in other words, make sure that your `pathToFiles` is the path on your local machine):

```{r}
pathToFiles = "./files/data/"

d1862 <- read.delim(paste0(pathToFiles, "dispatch_1862.tsv"), encoding="UTF-8", header=TRUE, quote="")
sw1 <- scan(paste0(pathToFiles, "sw1.md"), what="character", sep="\n")

```

The first file is articles from "The Daily Dispatch" for the year 1862. The newspaper was published in Richmond, VA --- the capital of the Confederate States (the South) during the American Civil War (1861-1865). The second file is a script of the first episode of Star Wars :).

### Libraries{#Lib07}

The following are the libraries that we will need for this section. Install those that you do not have yet.

```{r message=FALSE}
#install.packages("tidyverse", "readr", "stringr")
#install.packages("tidytext", "wordcloud", "RColorBrewer"", "quanteda", "readtext")

# General ones 
library(tidyverse)
library(readr)
library("RColorBrewer")

# Text Analysis Specific
library(stringr)
library(tidytext)
library(wordcloud)
library(quanteda)
library(readtext)
```
### Functions in `R` (a refresher){#functions07}

Functions are groups of related statements that perform a specific task, which help breaking a program into smaller and modular chunks. As programs grow larger and larger, functions make them more organized and manageable. Functions help avoiding repetition and makes code reusable.

Most programming languages, `R` including, come with a lot of pre-defined—or built-in—functions. Essentially, all statements that take arguments in parentheses are functions. For instance, in the code chunk above, `read.delim()` is a function that takes as its arguments: 1) filename (or, path to a file); 2) encoding; 3) specifies that the file has a header; and 4) not using `"` as a special character. We can also write our own functions, which take care of sets of operations thet we tend to repeat again and again. 

Later, take a look at this [video by one of the key `R` developers](https://campus.datacamp.com/courses/writing-functions-in-r/a-quick-refresher?ex=1), and check this [tutorial](https://rpubs.com/williamsurles/292234).

#### Simple Function Example: Hypothenuse{#Hypothenuse07}

(From [Wikipedia](https://en.wikipedia.org/wiki/Hypotenuse)) In geometry, a *hypotenuse* is the longest side of a right-angled triangle, the side opposite the right angle. The length of the hypotenuse of a right triangle can be found using the Pythagorean theorem, which states that the square of the length of the hypotenuse equals the sum of the squares of the lengths of the other two sides (*catheti*). For example, if one of the other sides has a length of 3 (when squared, 9) and the other has a length of 4 (when squared, 16), then their squares add up to 25. The length of the hypotenuse is the square root of 25, that is, 5.

Let's write a function that takes lengths of catheti as arguments and returns the length of hypothenuse:

```{r}
hypothenuse <- function(cathetus1, cathetus2) {
  hypothenuse<- sqrt(cathetus1*cathetus1+cathetus2*cathetus2)
  print(paste0("In the triangle with catheti of length ",
              cathetus1, " and ", cathetus2,
              ", the length of hypothenuse is ",
              hypothenuse))
  return(hypothenuse)
}
```


Let's try a simple example:

```{r}
hypothenuse(3,4)
```

Let's try a crazy example:

```{r}
hypothenuse(390,456)
```



## Homework{#HW07}

- Finish your worksheet and submit your HW as described below.
* *Additional:* if you'd like more practice, you can use `swirl` library:
  * To install: `install.packages("swirl")`
  * To run: `library(swirl)`
    * Then: `swirl()`
    * it will offer you a set of interactive exercises similar to DataCamp.

## Submitting homework{#SHW07}

* Homework assignment must be submitted by the beginning of the next class;
* Email your homework to the instructor as attachments.
	*  In the subject of your email, please, add the following: `57528-LXX-HW-YourLastName-YourMatriculationNumber`, where `LXX` is the number of the lesson for which you submit homework; `YourLastName` is your last name; and `YourMatriculationNumber` is your matriculation number.
	
**NB:** The original version has beed developed and prepared by Maxim Romanov for the course "R for Historical Research" (U Vienna, Spring 2019).