# Text Analysis II: Distances, Keywords, Summarization

```{r echo=FALSE}
library(knitr)
```

## Goals{#goals08}

- similarity distances; 
- keyword extraction (tf-idf);
- text summarization techniques;

## Preliminaries{#Prelim08}

### Data{#Data08}

```{r}
prep_fun = function(x) {
  x %>% 
    str_to_lower %>% # make text lower case
    str_replace_all("[^[:alnum:]]", " ") %>% # remove non-alphanumeric symbols
    str_replace_all("\\s+", " ") # collapse multiple spaces
}
```

```{r}
d1862 <- read.delim("./files/data/dispatch_1862.tsv", encoding="UTF-8", header=TRUE, quote="")
```

The following are the libraries that we will need for this section. Install those that you do not have yet.

```{r message=FALSE}
#install.packages("tidyverse", "readr", "stringr", "text2vec")
#install.packages("tidytext", "wordcloud", "RColorBrewer"", "quanteda", "readtext", "igraph")

# General ones 
library(tidyverse)
library(readr)
library("RColorBrewer")

# text analysis specific
library(stringr)
library(text2vec)
library(tidytext)
library(wordcloud)
library(quanteda)
library(readtext)
library(igraph)
```

## Document similarity/distance measures: `text2vec` library {#docsim08}

Document similarity—or distance—measures are valuable for a variety of tasks, such as identification of texts with similar (or the same) content. Let's just filter it down to some sample that would not take too much time to process. We also need to clean up our texts for better calculations.

### Distance Measures: Jaccard index, Cosine similarity, Euclidean distance 

The `text2vec` library can calculate a several different kinds of distances (details: <http://text2vec.org/similarity.html>): Jaccard, cosine, and Euclidean.

#### *Jaccard similarity/index*

is a simple measure of similarity based on the comparison of two sets, namely, as the proportion of the number of common words to the number of unique words in both documents. Jaccard similarity takes only unique set of words for each sentence/document (<https://en.wikipedia.org/wiki/Jaccard_index>). *Jaccard index* is commonly used to find text that deal with the same subjects (share same vocabulary --- frequencies of words have no effect on this measure)

Jaccard similarity measures the similarity between two nominal attributes by taking the intersection of both and divide it by their union.


#### *Cosine similarity*

another approach that measures similarity based on the content overlap between documents: each document is represented as a bag-of-words and as a sparse vector; the measure of overlap is defined as angle between vectors. *Cosine similarity* is better when we compare texts of varied length (angle of vectors, instead of distance).  (<https://en.wikipedia.org/wiki/Cosine_similarity>)

Cosine similarity measures the similarity between two vectors by taking the cosine of the angle the two vectors make in their dot product space. If the angle is zero, their similarity is one, the larger the angle is, the smaller their similarity. The measure is independent of vector length.

#### *Euclidean distance*

one of the most common measures---a straight-line distance between two points in Euclidian space; based on word frequencies and most commonly used to find duplicates (<https://en.wikipedia.org/wiki/Euclidean_distance>).

**NB**: more detailed explanations, see <https://cmry.github.io/notes/euclidean-v-cosine> 

#### Testing...

Let's try a small and simple example first.

```{r}

sentences = c("The Caliph arrived to Baghdad from Mecca.",
              "The Caliph arrived to Mecca from Baghdad.",
              "The Caliph arrived from Mecca to Baghdad. The Caliph arrived from Baghdad to Mecca.",
              "The Caliph arrived to Baghdad from Mecca. The Caliph arrived. The Caliph arrived. The Caliph arrived.",
              "The Caliph arrived to Baghdad from Mecca. The Caliph returned to Mecca from Baghdad.",
              "The Caliph arrived from Mecca to Baghdad, and then returned to Mecca.",
              "The vezier arrived from Isbahan to Mecca. The Caliph, Caliph, Caliph returned from Mecca to Baghdad Baghdad Baghdad.")

testDF <- data.frame("ID" = as.character(1:length(sentences)), "TEXT" = sentences)

testDF$TEXT <- prep_fun(testDF$TEXT)

```

Now, converting to `text2vec` format:

```{r}
# shared vector space
it = itoken(as.vector(testDF$TEXT))
v = create_vocabulary(it)
vectorizer = vocab_vectorizer(v)

# creating matrices
sparseMatrix = create_dtm(it, vectorizer)
denseMatrix = as.matrix(sparseMatrix)
```

Let's take a look inside:

```{r}
denseMatrix
```


```{r}
sparseMatrix
```

Let's generate our distance matrices:

```{r echo=F, message=FALSE, warning=FALSE}
jaccardMatrix = sim2(sparseMatrix, sparseMatrix, method = "jaccard", norm = "none")
cosineMatrix = sim2(sparseMatrix, sparseMatrix, method = "cosine", norm = "l2")
euclideanMatrix = dist2(denseMatrix, denseMatrix, method = "euclidean", norm="l2")
```

Now, let's check against the actual sentences:

```{r}
testDF$TEXT
```

For convenience, here they are again, in a more readable form:

1. The Caliph arrived to Baghdad from Mecca.
2. The Caliph arrived to Mecca from Baghdad.
3. The Caliph arrived from Mecca to Baghdad. The Caliph arrived from Baghdad to Mecca.
4. The Caliph arrived to Baghdad from Mecca. The Caliph arrived. The Caliph arrived. The Caliph arrived.
5. The Caliph arrived to Baghdad from Mecca. The Caliph returned to Mecca from Baghdad.
6. The Caliph arrived from Mecca to Baghdad, and then returned to Mecca.
7. The Vezier arrived from Isbahan to Mecca. The Caliph, Caliph, Caliph returned from Mecca to Baghdad Baghdad Baghdad.

```{r}
print("JACCARD: 1 is full match"); jaccardMatrix
print("COSINE: 1 is full match"); cosineMatrix
print("EUCLIDEAN: 0 is full match"); euclideanMatrix
```

All three distances tell us that 1, 2, and 3 are the "same". But when it comes to 4, the situation changes: Jaccard is most efficient, then Cosine, and Euclidean is least useful. If we want to find both 1 and 7, Cosine is the most effective, and Euclidean is the least effective.

Perhaps: Jaccard > overlap; Cosine > similarity; Euclidean > exactness?
Additional read: <https://cmry.github.io/notes/euclidean-v-cosine> (although `python` is used here)

### Now, let's run this on "Dispatch"

```{r}
sample_d1862 <- d1862 %>%
  filter(type=="advert")

sample_d1862$text <- prep_fun(sample_d1862$text)

# shared vector space
it = itoken(as.vector(sample_d1862$text))
v = create_vocabulary(it) %>%
  prune_vocabulary(term_count_min = 3) # 
vectorizer = vocab_vectorizer(v)
```

`prune_vocabulary()` is a useful function if you work with a large corpus; using `term_count_min=` would allow to remove low frequency vocabulary from our vector space and lighten up calculations.

Now, we need to create a document-feature matrix:

```{r}
dtmD = create_dtm(it, vectorizer)
```


```{r}
jaccardMatrix = sim2(dtmD, dtmD, method = "jaccard", norm = "none")
jaccardMatrix@Dimnames[[1]] <- as.vector(sample_d1862$id)
jaccardMatrix@Dimnames[[2]] <- as.vector(sample_d1862$id)
```

Let's take a look at a small section of our matrix. Can you read it? How should this data look in tidy format?
  
```{r}
jaccardMatrix[1:4, 1:2]
```

Converting matrix into a proper tidy data frame is a bit tricky. Luckily, `igraph` library can be extremely helpful here. We can treat our matrix as edges, where each number is the weight of each given edge. Loading this data into `igraph` will help us to avoid heavy-lifting on conversion as it can do all the complicated reconfiguration of our data, converting it into a proper dataframe that conforms to the principles of tidy data. 

All steps include:
  
1) convert our initial object from a *sparse* matrix format into a *regular* matrix format;
2) rename rows and columns (we have done this already though);
3) create `igraph` object from our regular matrix;
4) extract edges dataframe.

```{r message=FALSE}
jaccardMatrix <- as.matrix(jaccardMatrix)

library(igraph)
jaccardNW <- graph.adjacency(jaccardMatrix, mode="undirected", weighted=TRUE)
jaccardNW <- simplify(jaccardNW)
jaccard_sim_df <- as_data_frame(jaccardNW, what="edges")

colnames(jaccard_sim_df) <- c("text1", "text2", "jaccardSimilarity")

jaccard_sim_df <- jaccard_sim_df %>%
  arrange(desc(jaccardSimilarity))

head(jaccard_sim_df, 10)
```

```{r}
t_jaccard_sim_df_subset <- jaccard_sim_df %>%
  filter(jaccardSimilarity > 0.49) %>%
  filter(jaccardSimilarity <= 0.9) %>%
  arrange(desc(jaccardSimilarity), .by_group=T)

head(t_jaccard_sim_df_subset, 10)
```

Let's check the texts of `1862-04-07_advert_175` and `1862-04-05_advert_38`, which have the score of 0.9000000 (a close match).

```{r}
example <- d1862 %>%
  filter(id=="1862-04-07_advert_175")
```

```
[1] "Very desirable Residence on the South of Main st., between and Cheery
streets, in Sidney, at Auction. -- We will sell, upon the premises, on
Monday, the 7th day of April, at 4½ o'clock P. M., a very comfortable and
well arranged Framed Residence located as above, and now in the occupancy
of Mr. Wm. B Davidson It 7 rooms with closed, kitchen and all accessary out
building, and is particularly adapted for the accommodation of a medium
family. The location of this house is as desirable as any in Sidney; is
located in a very pleasant neighborhood, within a few minutes walk of the
business portion of the city. The lot fronts 30 feet and runs back 189 feet
to an alley 30 feet wide. Terms. -- One-third cash; the balance at 6 and
12 months, for negotiable notes, with interest added, and secured by a
trust deed. The purchaser to pay the taxes and insurance for 1862. Jas.
M. Taylor, & Son, Auctioneers. mh 27"
```


```{r}
example <- d1862 %>%
  filter(id=="1862-04-05_advert_38")
```

```
[1] "Very desirable Framed Residence of the South side of Main St. Between
culvert and Cherri streets. In Sidney, at Auction. -- We will sell, upon
the premises, on Monday, the 7th day of April, at 4½ o'clock P. M. a very
comfortable and well arranged Framed. Residence located as above, and now
in the occupancy of Mr. Wm. B Davidson. It contains 7 rooms, with closets,
kitchen and all necessary out buildings, and is particularly adapted for
the accommodation of a medium sized family. The location of this house is
as desirable as any in Sidney; is located in a very pleasant neighborhood,
and within a few minutes walk of the business portion of the city. The lot
fronts 80 feet and runs back 189 feet to an alley 20 feet wide. Terms. --
One-third cash, the balance at 6 and 12 months, for negotiable notes, with
interest added, and secured by a trust deed. The purchaser to pay the taxes
and insurance for 1862 Jas. M. Taylor & Son. mh 27 Auctioneers."
```

(@) Check <http://text2vec.org/similarity.html> and calculate `cosine` and `euclidean` distances for the same set of texts. What is the score for the same two texts? How do these scores differ in your opinion?

>> your observations


<!--
```{r}
# FOR YOU TO KEEP IN MIND

dtmD = create_dtm(it, vectorizer)
dtmD2 <- as.matrix(dtmD)
euclideanMatrix = dist2(dtmD2, dtmD2, method = "euclidean", norm = "l2")
colnames(euclideanMatrix) <- as.vector(sample_d1862$id)
rownames(euclideanMatrix) <- as.vector(sample_d1862$id)

# names to rows and columns in dense/regular matrices are assigned differently.

```
-->

(@) Choose one of the distance measures and take a close look at a subset of texts with the closest match (i.e. find a text which has the highest number of complete matches --- 1.0). Try to apply as many techniques as possible in your analysis (e.g., frequency lists, wordclouds, graphing over time, etc.)

>> your analysis, your code...

## TF-IDF

Before we proceed, let's load some text. Below is an example of how you can load a text using its URL. However, be mindful about using this approach: it is convenient with a small number of short texts, but not efficient with large number of long texts.


```{r}
urlUDHR <- "https://univie-histr-2020s.github.io/files/UDHR.csv"
udhr <- read.delim(url(urlUDHR), encoding="UTF-8", header=TRUE, quote="", stringsAsFactors = FALSE)
```


```{r}
udhrTidy <- udhr %>%
  unnest_tokens(WORD, TEXT) %>%
  count(SECTION, WORD, sort=TRUE)

summary(udhrTidy)
```

**From Wikipedia**: In information retrieval, tf–idf or TFIDF, short for term frequency–inverse document frequency, is a numerical statistic that is intended to reflect how important a word is to a document in a collection or corpus.[1] It is often used as a weighting factor in searches of information retrieval, text mining, and user modeling. The tf–idf value increases proportionally to the number of times a word appears in the document and is offset by the number of documents in the corpus that contain the word, which helps to adjust for the fact that some words appear more frequently in general. tf–idf is one of the most popular term-weighting schemes today; 83% of text-based recommender systems in digital libraries use tf–idf. Variations of the tf–idf weighting scheme are often used by search engines as a central tool in scoring and ranking a document's relevance given a user query. tf–idf can be successfully used for stop-words filtering in various subject fields, including text summarization and classification. One of the simplest ranking functions is computed by summing the tf–idf for each query term; many more sophisticated ranking functions are variants of this simple model.

```{r}
udhr_TFIDF <- udhrTidy %>% 
  bind_tf_idf(WORD, SECTION, n) %>%
  arrange(desc(tf_idf)) %>%
  ungroup

udhr_TFIDF %>%
  filter(tf_idf >= 0.15)
```


```{r}
hist(udhr_TFIDF$tf_idf)
```

Let's take a look at any of the Articles:

```{r}
articleID = "Article 4"
temp <- filter(udhr_TFIDF, SECTION==articleID) %>%
  arrange(desc(tf_idf))

temp
```

```{r}
articleID = "Article 26"
temp <- filter(udhr_TFIDF, SECTION==articleID) %>%
  arrange(desc(tf_idf))

temp
```

We can use wordcloud to vizualize results --- but they will not be too telling, if we use word frequencies.

```{r message=F, fig.height=10, fig.width=10}
library(wordcloud)
library("RColorBrewer")

set.seed(1234)
wordcloud(words=temp$WORD, freq=temp$n,
          min.freq = 1, rot.per = .0, random.order=FALSE, scale=c(10,.5),
          max.words=150, colors=brewer.pal(8, "Dark2"))
```

Instead we can use `tf_idf` values:

```{r message=F, fig.height=10, fig.width=10}
set.seed(1234)
wordcloud(words=temp$WORD, freq=temp$tf_idf,
          min.freq = 1, rot.per = .0, random.order=FALSE, scale=c(10,.5),
          max.words=150, colors=brewer.pal(8, "Dark2"))
```

### Inaugural speeches of the US presidents

The `quanteda` package includes a corpus of presidential inaugural speeches. What did the presidenst speak about? For thoughts and ideas, take a look at what Google News Lab did with this data: <http://inauguratespeeches.com/>. You can find readable addresses here: <https://www.presidency.ucsb.edu/documents/presidential-documents-archive-guidebook/inaugural-addresses>.

```{r}
data("data_corpus_inaugural", package = "quanteda")
inaug_dfm <- quanteda::dfm(data_corpus_inaugural, verbose = FALSE)
head(inaug_dfm)
```

The corpus is stored as a `document-feature matrix`, which we can convert into a more familiar tidy format in the following manner:

```{r}
inaug_td <- tidy(inaug_dfm)
head(inaug_td)
```

(@) Using what you have learned so far, analyze speeches of American presidents:

>> your code; you analysis; your observations...

## Text summarization

Before we proceed, let's load some text. Below is an example of how you can load a text using its URL. However, be mindful about using this approach: it is convenient with a small number of short texts, but not efficient with large number of long texts.

```{r}
urlText <- "https://univie-histr-2020s.github.io/files/test_text.txt"
testText <- scan(url(urlText), what="character", sep="\n")
```

We can use different algorithms to *summarize* texts, which in this context means extracting key sentences, whose keyness is calculated through different means. Library `lexRankr` used tf-idf values and some methods from social network analysis to identify the most central sentences in a text.  (For technical details, see: <https://cran.r-project.org/web/packages/lexRankr/lexRankr.pdf>; for a detailed *math* description: <http://www.cs.cmu.edu/afs/cs/project/jair/pub/volume22/erkan04a-html/erkan04a.html>).

Take a look at the summary and then at the full text!

```{r message=FALSE, warning=FALSE}
library(lexRankr)
```


```{r}
textToSummarize = testText

summary = lexRank(textToSummarize,
                  docId = rep(1, length(textToSummarize)), #repeat same docid for all of input vector
                  n = 5, # number of sentences
                  continuous = TRUE)

# this is just preparing results for better viewing
summary$sentenceId <- str_replace_all(summary$sentenceId, "\\d+_", "")
summary$sentenceId <- as.numeric(summary$sentenceId)

summary <- summary %>%
  arrange(sentenceId)

summary$sentence
```

```
[1] "The Klimt University of Vienna Ceiling Paintings, also known
as the Faculty Paintings, were a series of paintings made by Gustav
Klimt for the ceiling of the University of Vienna`s Great Hall
between the years of 1900–1907."

[2] "Upon presenting his paintings, Philosophy, Medicine and
Jurisprudence, Klimt came under attack for `pornography` and
`perverted excess` in the paintings."                                                                                                                  
[3] "Klimt described the painting as follows: `On the left a
group of figures, the beginning of life, fruition, decay."                                                                                                                                                          
[4] "In 1903, Hermann Bahr, a writer and a supporter of Klimt, in
response to the criticism of the Faculty Paintings compiled articles
which attacked Klimt, and published a book Gegen Klimt (Against Klimt)
with his foreword, where he argued that the reactions were absurd."

[5] "In 1911 Medicine and Jurisprudence were bought by Klimt`s
friend and fellow artist, Koloman Moser.[clarification needed]
Medicine eventually came into the possession of a Jewish family,
and in 1938 the painting was seized by Germany." 
```

Here is another corpus of articles to play with (from `tm` library):

```{r message=FALSE, warning=FALSE}
library(tm)

data("acq")

acqTidy <- tidy(acq)

```

Here is an article:

```{r}
item = 1

test <- str_replace_all(acqTidy$text[item], "\\s+", " ")
test
```
```
[1] "Computer Terminal Systems Inc said it has completed the sale
of 200,000 shares of its common stock, and warrants to acquire an
additional one mln shares, to <Sedio N.V.> of Lugano, Switzerland
for 50,000 dlrs. The company said the warrants are exercisable for
five years at a purchase price of .125 dlrs per share. Computer
Terminal said Sedio also has the right to buy additional shares
and increase its total holdings up to 40 pct of the Computer
Terminal's outstanding common stock under certain circumstances
involving change of control at the company. The company said if the
conditions occur the warrants would be exercisable at a price equal
to 75 pct of its common stock's market price at the time, not to
exceed 1.50 dlrs per share. Computer Terminal also said it sold the
technolgy rights to its Dot Matrix impact technology, including any
future improvements, to <Woodco Inc> of Houston, Tex. for 200,000
dlrs. But, it said it would continue to be the exclusive worldwide
licensee of the technology for Woodco. The company said the moves
were part of its reorganization plan and would help pay current
operation costs and ensure product delivery. Computer Terminal makes
computer generated labels, forms, tags and ticket printers and
terminals. Reuter"
```

Let's see how our summary comes out:

```{r}
textToSummarize = test

summary = lexRank(textToSummarize,
                  docId = rep(1, length(textToSummarize)), #repeat same docid for all of input vector
                  n = 3, # number of sentences
                  continuous = TRUE)

# this is just preparing results for better viewing
summary$sentenceId <- str_replace_all(summary$sentenceId, "\\d+_", "")
summary$sentenceId <- as.numeric(summary$sentenceId)

summary <- summary %>%
  arrange(sentenceId)

summary$sentence

```
```
[1] "Computer Terminal Systems Inc said it has completed the
sale of 200,000 shares of its common stock, and warrants to acquire
an additional one mln shares, to <Sedio N.V.> of Lugano,
Switzerland for 50,000 dlrs."

[2] "Computer Terminal said Sedio also has the right to buy
additional shares and increase its total holdings up to 40 pct
of the Computer Terminal's outstanding common stock under certain
circumstances involving change of control at the company."

[3] "The company said if the conditions occur the warrants
would be exercisable at a price equal to 75 pct of its common
stock's market price at the time, not to exceed 1.50 dlrs per share."
```


(@) Try this with other articles --- and compare summaries with full versions. Share your observations.

>> your code; your observations


## Homework{#HW08}

* Read about *ngrams* in *Chapter 4. Relationships between words: n-grams and correlations* (<https://www.tidytextmining.com/ngrams.html>), in <https://www.tidytextmining.com/>.
  - Using what you have learned in this chapter identify and analyze *bigrams* in Dispatch, 1862.
  - Submit the results of your analysis as an R notebook, as usual.
  - You are welcome to work in groups.
* *Optional*: Work through Chapter 9 of Arnold, Taylor, and Lauren Tilton. 2015. *Humanities Data in R*. New York, NY: Springer Science+Business Media. (on Moodle!): create a notebook with all the code discusses there and send it via email (share via DropBox or some other service, if too large).
* DataCamp Assignments.

## Submitting homework{#SHW08}

* Homework assignment must be submitted by the beginning of the next class;
* Email your homework to the instructor as attachments.
	*  In the subject of your email, please, add the following: `57528-LXX-HW-YourLastName-YourMatriculationNumber`, where `LXX` is the number of the lesson for which you submit homework; `YourLastName` is your last name; and `YourMatriculationNumber` is your matriculation number.