# Text Analysis III: Finding Groups of Texts

```{r echo=FALSE}
library(knitr)
```

## Goals{#goals09}

- introduce text clustering approaches: approaches that allow us find groups of thematically similar texts:
  - hierarchical clustering; 
  - k-means clustering;
  - topic modeling;

## Preliminaries{#Prelim09}

### Libraries

The following are the libraries that we will need for this section. Install those that you do not have yet.

```{r message=FALSE}
#install.packages("ggplot2", "LDAvis", "readr", "slam", "stringr", "tictoc", "tidytext", "tidyverse", "tm", "topicmodels")

# general
library(ggplot2)

# text analysis specific
library(readr)
library(slam)
library(stringr)
library(tidytext)
library(tidyverse)
library(tm)
library(topicmodels)

library(text2vec)
library(stylo)

library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization

# extra
library(tictoc) # to time operations
```

#### The Dispatch Data: Preprocessing

- [The Richmond Dispatch (1861)](./files/data/dispatch_1861.tsv)
- [The Richmond Dispatch (1862)](./files/data/dispatch_1862.tsv)
- [The Richmond Dispatch (1863)](./files/data/dispatch_1863.tsv)
- [The Richmond Dispatch (1864)](./files/data/dispatch_1864.tsv)

We will limit to only one year, but, of course, the results are always more interesting with more data...

Loading...

```{r}
d1864 <- read.delim("./files/data/dispatch_1864.tsv", encoding="UTF-8", header=TRUE, quote="", stringsAsFactors = FALSE)
d1864$date <- as.Date(d1864$date, format="%Y-%m-%d")
```


```{r}
knitr::kable(head(d1864), table.attr = "class=\"striped\"", format="html")
```

...


## Homework{#HW09}

* given in the chapter.

## Submitting homework{#SHW09}

* Homework assignment must be submitted by the beginning of the next class;
* Email your homework to the instructor as attachments.
	*  In the subject of your email, please, add the following: `57528-LXX-HW-YourLastName-YourMatriculationNumber`, where `LXX` is the number of the lesson for which you submit homework; `YourLastName` is your last name; and `YourMatriculationNumber` is your matriculation number.