# Text Analysis I: Basics

## ...

## Goals:{#goals07}

- introduce basic text analysis approaches

## Preliminaries{#Prelim07}

### Data{#Data07}

We will use the following text files in this worksheet. Please download them and keep them close to your worksheet. Since some of the files are quite large, you want to download them before loading them into R:

* [The Richmond Dispatch (1862)](./files/data/dispatch_1862.tsv)
* [Star Wars I: The Phantom Menace (script)](./files/data/sw1.md)

In order to make loading these files a little bit easier, you can paste the path to where you placed these files into an isolated variable and then reuse it as follows (in other words, make sure that your `pathToFiles` is the path on your local machine):

```{r}
pathToFiles = "./files/data/"

d1862 <- read.delim(paste0(pathToFiles, "dispatch_1862.tsv"), encoding="UTF-8", header=TRUE, quote="")

sw1 <- scan(paste0(pathToFiles, "sw1.md"), what="character", sep="\n")

```

The first file is articles from "The Daily Dispatch" for the year 1862. The newspaper was published in Richmond, VA --- the capital of the Confederate States (the South) during the American Civil War (1861-1865). The second file is a script of the first episode of Star Wars :).

### Libraries{#Lib07}

The following are the libraries that we will need for this section. Install those that you do not have yet.

```{r message=FALSE}
#install.packages("tidyverse", "readr", "stringr")
#install.packages("tidytext", "wordcloud", "RColorBrewer"", "quanteda", "readtext")

# General ones 
library(tidyverse)
library(readr)
library("RColorBrewer")

# Text Analysis Specific
library(stringr)
library(tidytext)
library(wordcloud)
library(quanteda)
library(readtext)
```



## Homework{#HW07}

- Finish your worksheet and submit your HW as described below.
* *Additional:* if you'd like more practice, you can use `swirl` library:
  * To install: `install.packages("swirl")`
  * To run: `library(swirl)`
    * Then: `swirl()`
    * it will offer you a set of interactive exercises similar to DataCamp.

## Submitting homework{#SHW07}

* Homework assignment must be submitted by the beginning of the next class;
* Email your homework to the instructor as attachments.
	*  In the subject of your email, please, add the following: `57528-LXX-HW-YourLastName-YourMatriculationNumber`, where `LXX` is the number of the lesson for which you submit homework; `YourLastName` is your last name; and `YourMatriculationNumber` is your matriculation number.
	
**NB:** The original version has beed developed and prepared by Maxim Romanov for the course "R for Historical Research" (U Vienna, Spring 2019).